---
title: "DeltaOutflowPlotting"
author: "Caroline Newell"
date: "2/23/2022"
output: html_document
---
```{r}
library(readr)
library(tidyverse)
library(lubridate)
library(ggplot2)
library(lfstat)
```

```{r}
DeltaFlow_WaterYears1960_2021 <- read_csv("1960 to present/data/DeltaFlow_WaterYears1960_2021.csv")

drought_years<-c(1976, 1977, 1987, 1988, 1989, 1990, 1991, 1992, 2007, 2008, 2009, 2012, 2013, 2014, 2015, 2016, 2020, 2021) #data from: https://water.ca.gov/water-basics/drought  and Peter (2020, 2021)

```

# Now we plot
```{r}
#basic bubble plots
ggplot(DeltaFlow1960_2021, aes(x=Date, y=TOT)) +
  geom_line() +
  xlab("")+
  ylab("Inflow")

ggplot(DeltaFlow1960_2021, aes(x=Date, y=EXPORTS)) +
  geom_line() +
  xlab("")+
  ylab("Exports")

ggplot(DeltaFlow1960_2021, aes(x=Date)) +
  geom_line(aes(y=TOT), color = "steelblue") +
  geom_line(aes(y=EXPORTS), color = "darkred", linetype = "twodash") +
  xlab("Water Year") +
  theme(legend.position = c(.95, .95),
        legend.justification = c("right", "top"),
        legend.box.just = "right",
        legend.margin = margin(6, 6, 6, 6))

#Wont make a legend :( Gotta collapse the two variables into key-value pairs and then visualize

df<- DeltaFlow1960_2021 %>% 
  select(Date, TOT, EXPORTS) %>% 
  gather(key="variable", value = "value", -Date)
head(df)

ggplot(df, aes(x=Date, y=value))+
  geom_line(aes(color = variable)) +
  scale_color_manual(values = c("darkred", "steelblue"))+
  facet_wrap(~variable, scales = "free")
```

John's response: You are close! There is so much noise that it makes it hard to read a signal. Try summing up by years, then create a stacked plot on one axis in MAF and see if we can get some traction.

```{r}
ggplot(DeltaFlow1960_2021_WaterYear, aes(x=WaterYear))+
  geom_bar(aes(y=TOT_MAF), stat="identity", color="black", fill="steelblue")+
  geom_line(aes(y=Exports_MAF, group =1), color="red", size=1.1) +
  xlab("Water Year") + ylab("MAF")

str(DeltaFlow1960_2021_WaterYear)
#Export trend line:
ggplot(DeltaFlow1960_2021_WaterYear, aes(x=WaterYear)) +geom_rect(aes(xmin=1976, xmax=1977, ymin=0, ymax=Inf), fill = "grey75", alpha = .1) +
  geom_rect(aes(xmin=1987, xmax=1992, ymin=0, ymax=Inf),fill = "grey75", alpha = .1) +
  geom_rect(aes(xmin=2007, xmax=2009, ymin=0, ymax=Inf),fill = "grey75", alpha = .1) +
  geom_rect(aes(xmin=2012, xmax=2016, ymin=0, ymax=Inf),fill = "grey75", alpha = .1)+
  geom_bar(aes(y=TOT_MAF), stat="identity", color="black", fill="steelblue") + 
  geom_line(aes(y=Exports_MAF, group =1), color="red", size=1.1) + xlab("Water Year") +  ylab("MAF") + ggtitle("Total Delta Inflow and Exports")+ scale_x_continuous(breaks=c(1960, 1970, 1980, 1990, 2000, 2010, 2020))+
  theme_classic(base_size=16) 

#Export bars: 
ggplot(DeltaFlow1960_2021_WaterYear, aes(x=WaterYear)) +geom_rect(aes(xmin=1976, xmax=1977, ymin=0, ymax=Inf), fill = "grey75", alpha = .1) +
  geom_rect(aes(xmin=1987, xmax=1992, ymin=0, ymax=Inf),fill = "grey75", alpha = .1) +
  geom_rect(aes(xmin=2007, xmax=2009, ymin=0, ymax=Inf),fill = "grey75", alpha = .1) +
  geom_rect(aes(xmin=2012, xmax=2016, ymin=0, ymax=Inf),fill = "grey75", alpha = .1)+
  geom_rect(aes(xmin=2020, xmax=2021, ymin=0, ymax=Inf),fill = "grey75", alpha = .1)+
  geom_bar(aes(y=TOT_MAF), stat="identity", color="black", fill="steelblue") + 
  geom_bar(aes(y=Exports_MAF), stat="identity", color="black", fill="red") + xlab("Water Year") +  ylab("MAF") + ggtitle("Total Delta Inflow and Exports")+ scale_x_continuous(breaks=c(1960, 1970, 1980, 1990, 2000, 2010, 2020))+
  theme_classic(base_size=16) 
```

John's response:
Thanks for checking the effect of WY. I'm reasonable confident that the line+circles on the old plot is Delta Inflow (or QTOT). Could the blue + red colors be export across the two pumping facilities?

Anyway, I prefer your plot. Would you like to send these to Peter with a short explanation? Send one with the trend line and one with the stacked bars for Export. And could you use 10 yr increments on the x axis?
Nicely done. 
For fun, will you try one more thing? Using the stacked bar plot, can you plot the mean annual location of X2 across the time series? For this, you will need a second axis. If you get stuck, let me know.