---
title: "Exports and Inflow to the Delta"
author: "Caroline Newell"
date: "2/1/2022"
output: word_document
editor_options: 
  chunk_output_type: console
---
Hi Caroline, I have a small task for you. Could you generate a graph of Exports and Inflow to the Delta using data from this website? https://www.usbr.gov/mp/cvo/pmdoc.html
Attached is a version of the kind of plot we would like--we just need an update. It doesn't have to look exactly like this by any means, we just need to show the relationship. Also attached is a spreadsheet showing last year's values (downloaded from that website--we just need the longer time series, dating from the 60s, and a nice plot.  Thank you!
john

John's website sucked. Better option: https://data.cnra.ca.gov/dataset/dayflow
```{r}
library(readr)
library(tidyverse)
library(lubridate)
library(ggplot2)
library(lfstat)
#install_github("ChandlerLutz/ggts")
```

#Reading al of the data in, formatting, writing

```{r}
dayflow_results_1956_1969 <- read_csv("C:/Users/cnewe/OneDrive/Documents/Delta Inflow Outflow/1960 to present/data/dayflow-results-1956-1969.csv")
dayflow_results_1970_1983 <- read_csv("C:/Users/cnewe/OneDrive/Documents/Delta Inflow Outflow/1960 to present/data/dayflow-results-1970-1983.csv")
dayflow_results_1984_1996 <- read_csv("C:/Users/cnewe/OneDrive/Documents/Delta Inflow Outflow/1960 to present/data/dayflow-results-1984-1996.csv")
dayflow_results_1997_2020 <- read_csv("C:/Users/cnewe/OneDrive/Documents/Delta Inflow Outflow/1960 to present/data/dayflow-results-1997-2020.csv")
dayflowcalculations2021 <- read_csv("C:/Users/cnewe/OneDrive/Documents/Delta Inflow Outflow/1960 to present/data/dayflowcalculations2021.csv")
```

Now I need to join all of the data together into one dataframe.

```{r}
#str(dayflowcalculations2021$Date)
#dayflowcalculations2021$Date<-mdy(dayflowcalculations2021$Date) #This deletes values for some reason, and date data appears to be correct already... so i'm gonna leave it.
                                               dayflow_results_1997_2020$Date<-mdy(dayflow_results_1997_2020$Date)

#str(dayflow_results_1984_1996) #Date is saved as a character, but when i do as.date it changes the value completely...                                                                                             
dayflow_results_1984_1996$Date<-mdy(dayflow_results_1984_1996$Date)
#str(dayflow_results_1984_1996$Date)
                                                  dayflow_results_1970_1983$Date<- mdy(dayflow_results_1970_1983$Date)
                                                    dayflow_results_1956_1969$Date<- mdy(dayflow_results_1956_1969$Date)
                                                    
#Need to pull out only relevant columns since all columns don't match between data sets. 

dayflowcalculations2021_subset<- dayflowcalculations2021[ , c('Date', 'TOT', 'EXPORTS', 'OUT')]  #how to subset with base r                                                  
dayflow_results_1997_2020_subset<-dayflow_results_1997_2020 %>% select(Date, TOT, EXPORTS, OUT) #how to subset using tidyverse

dayflow_results_1984_1996_subset<-dayflow_results_1984_1996 %>% select(Date, TOT, EXPORT, OUT)
#need to rename EXPORT -> EXPORTS
dayflow_results_1984_1996_subset<-dayflow_results_1984_1996_subset %>% rename(EXPORTS = EXPORT)
str(dayflow_results_1984_1996_subset)

dayflow_results_1970_1983_subset<-dayflow_results_1970_1983 %>% select(Date, TOT, EXPORT, OUT)
#need to rename EXPORT -> EXPORTS
dayflow_results_1970_1983_subset<- dayflow_results_1970_1983_subset %>% rename(EXPORTS = EXPORT)
str(dayflow_results_1970_1983_subset)

dayflow_results_1956_1969_subset<- dayflow_results_1956_1969 %>% select(Date, TOT, EXPORT, OUT)
#need to rename EXPORT -> EXPORTS
dayflow_results_1956_1969_subset<-dayflow_results_1956_1969_subset %>% rename(EXPORTS = EXPORT)

#okay, now I think we are ready to bind!

DeltaFlow1956_2021<- bind_rows(dayflow_results_1956_1969_subset, dayflow_results_1970_1983_subset, dayflow_results_1984_1996_subset, dayflow_results_1997_2020_subset, dayflowcalculations2021_subset)

#View(DeltaFlow1956_2021) #Dopenessss
str(DeltaFlow1956_2021)
```

```{r}
#Now I only want WATER YEAR 1960 through 2021.
DeltaFlow1960_2021<- DeltaFlow1956_2021 %>% filter(Date >="1959-10-01")
#View(DeltaFlow1960_2021)
#Save that shit!
write.csv(DeltaFlow1960_2021, "C:/Users/cnewe/OneDrive/Documents/Delta Inflow Outflow/1960 to present/data/DeltaFlow_WaterYears1960_2021.csv")
```

